---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 S26 Assignment 5"
author: "Tate Whittaker"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
```

## Problem 1
Convert the following to date or date/time objects.

a. September 13, 2010.

```{r}
mdy("September 13, 2010")
```

b. Sept 13, 2010.

```{r}
mdy("Sept 13, 2010")
```

Here I get a warning because "Sept" isn't an appropriate month for the mdy function.

c. Sep 13, 2010.

```{r}
mdy("Sep 13, 2010")
```
d. S 13, 2010. Comment on the month abbreviation needs.

```{r}
mdy("S 13, 2010")
```

Here I get a warning because "S" isn't an appropriate month for the mdy function,
similar to Sept..

e. 07-Dec-1941.

```{r}
dmy("07-Dec-1941")
```

f. 1-5-1998. Comment on why you might be wrong.

```{r}
mdy("1-5-1998")
```

I might be wrong here because 01 could be the month or the day, and same thing with
05, it could also be the month or the day.

g. 21-5-1998. Comment on why you know you are correct.

```{r}
dmy("21-5-1998")
```

I know am I right here because the only possibiliy for month is 05, meaning 21 
has to be the day.

h. 2020-May-5 10:30 am

```{r}
ymd_hm("2020-May-5 10:30 am")
```

i. 2020-May-5 10:30 am PDT (ex Seattle)

```{r}
ymd_hm("2020-May-5 10:30 am", tz="America/Vancouver")
```

j. 2020-May-5 10:30 am AST (ex Puerto Rico)

```{r}
ymd_hm("2020-May-5 10:30 am", tz="America/Puerto_Rico")
```

## Problem 2

Using just your date of birth (ex Sep 7, 1998) and today's date calculate the following:

```{r}
tate_birthday <- mdy("Mar 24, 2004")
```

a. Calculate the date of your 64th birthday.

```{r}
(tate_birthday + years(64))
```

b. Calculate your current age (in years).

```{r}
current_age <- year(as.period(interval(tate_birthday, Sys.Date())))
current_age
```

c. Using your result in part (b), calculate the date of your next birthday.

```{r}
next_birthday <- years(current_age) + tate_birthday + years(1)
next_birthday
```

d. The number of _days_ until your next birthday.

```{r}
day_interval <- interval(Sys.Date(), next_birthday)
as.period(day_interval, units = "days")
```

e. The number of _months_ and _days_ until your next birthday.

```{r}
as.period(day_interval, units = "months")
```

## Problem 3

Suppose you have arranged for a phone call to be at 3 pm on May 8, 2015 at Arizona time. However, the recipient will be in Auckland, NZ. What time will it be there? 

```{r}
az_time <- mdy_h("May 8, 2015 3 pm", tz = "US/Arizona")
```

```{r}
with_tz(az_time, tz="Pacific/Auckland")
```

The time in Auckland, NZ would be 10:00:00 on May 9th, 2015.

## Problem 4

It turns out there is some interesting periodicity regarding the number of births on particular days of the year.

a. Using the `mosaicData` package, load the data set `Births78` which records the number of children born on each day in the United States in 1978. Because this problem is intended to show how to calculate the information using the `date`, remove all the columns *except* `date` and `births`. 

```{r}
library(mosaicData)
data(Births78)
Births78 <- Births78 %>%
  select(date, births)
head(Births78)
```

b. Graph the number of `births` vs the `date` with date on the x-axis. What stands out to you? Why do you think we have this trend?

```{r}
births_graph <- ggplot( Births78, aes(x = date, y = births) ) +
  geom_point()
births_graph
```

The graph points follow a wave pattern, with high births in January of 1978, then
a dip in the spring of 1978, followed by an increase in the fall and winter of 1978.
This could be because of weather, geo-political reasons, or holidays.

c. To test your assumption, we need to figure out the what day of the week each observation is. Use `dplyr::mutate` to add a new column named `dow` that is the day of the week (Monday, Tuesday, etc). This calculation will involve some function in the `lubridate` package and the `date` column.

```{r}
Births78 <- Births78 %>%
  mutate(
    dow = wday( date, label = TRUE, abbr = FALSE )
  )
```

d. Plot the data with the point color being determined by the day of the week variable.

```{r}
births_weeks_graph <- ggplot( Births78, aes(x = date, y = births) ) +
  geom_point(aes( color = dow ) )
births_weeks_graph
```










