---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 Assignment 1"
author: "Tate Whittaker"
date: "2026-2-17"
output: pdf_document
---

# Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`. Print only the result `vec_c`. Do not use the print command.

```{r}
vec_a <- c(2, 4, 6)
vec_b <- c(8, 10, 12)
(vec_c <- vec_a + vec_b)
```

# Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a` and show the result. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?

```{r}
vec_d <- c(14, 20)
(vec_a + vec_d)
```

The result is 16, 24, 20 and R recylced vec_d, meaning it created a hidden vector
that looked like (14, 20, 14) and added it to vec_a. The warning that R gave
basically said that the longer object length (vec_a) was not a multiple of the 
shorter object length (vec_d), meaning that 3 is not a multiple of 2.

# Problem 3

Next add 5 to the vector `vec_a` and show the result. What is the result and what did R do? Why doesn't R give you a warning message similar to what you saw in the previous problem?

```{r}
(vec_a + 5)
```

Again, R created a hidden structure that looked like (5, 5, 5) and added it to
vec_a. This time R didn't give a warning message similar to Problem 2 because the
longer object length (vec_a) is a mutliple of the shorter object length.

# Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function

```{r}
seq(1, 5, 1)
```

b. Using the `a:b` shortcut.

```{r}
1:5
```

# Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the `seq()` function  

```{r}
seq(2, 20, 2)
```

b. Using the a:b shortcut and some subsequent algebra. 

```{r}
1:10 * 2
```

# Problem 6

Generate a vector of 21 elements that are evenly spaced between 0 and 1 using the `seq()` command and name this vector `x`. Print the vector `x`.

```{r}
(x <- seq(0, 1, length.out = 21))
```

# Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep( c(2, 4, 8), 3)
```

# Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for `rep()` to see all of the options that `rep()` will accept. In particular, look at the optional argument `each=`.

```{r}
rep( c(2, 4, 8), each = 4)
```
  
# Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Follow the prompts below to create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands. Print `M`.

```{r}
(M <- matrix( seq(2, 30, 2), nrow = 3, ncol = 5, byrow = TRUE ))
```

ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command. Print `M`

```{r}
row1 <- seq(2, 10, 2)
row2 <- seq(12, 20, 2)
row3 <- seq(22, 30, 2)
(M <- rbind(row1, row2, row3))
```

b. Extract the second row out of `M`.

```{r}
M[2,]
```

c. Extract the element in the third row and second column of `M`.

```{r}
M[3,2]
```

# Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```

The first method removes all indexes in the data frame that have an 
NA value in Grades.
The second method keeps/shows all indexes in the data fram that do not have an 
NA value in Grades.

# Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
        
```{r}
x <- c(4,5,6,7,8,9,10)
y <- c(34,35,41,40,45,47,51)
slope <- 2.82
p.value <- 0.000131

my.test <- list(X = x, Y = y, Slope = slope, P_Value = p.value)
```

To show you have created your list correctly, type use `str()` to show the structure of the list.

b.  Extract the second element in the list.

```{r}
my.test[2]
```

c.  Extract the element named `p.value` from the list.

```{r}
my.test["P_Value"]
```



